---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aRm

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/ddebeer/aRm/graph/badge.svg)](https://app.codecov.io/gh/ddebeer/aRm)
[![R-CMD-check](https://github.com/ddebeer/aRm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ddebeer/aRm/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/aRm)](https://CRAN.R-project.org/package=aRm)
<!-- badges: end -->

The aRm packages provides helpful functions related to the course Analysis of Repeated Measurements

## Installation

You can install the development version of aRm from [GitHub](https://github.com/) with:

```{r install aRm}
#| eval: false
# install.packages("pak")
pak::pak("ddebeer/aRm")
```

## Example - Linear Hypothesis

This is a basic example which shows how to use the `extract_results()` function.

```{r example}
library(aRm)
```


For this example, let's load the OBrienKaiser data. The data set includes 15 repeated measures: 

  - five pre measures 
  - five post measures
  - five follow-up measures

```{r load data}
data(OBrienKaiser, package = "carData")
```

For more information about the data set, you can consult the documentation.

```{r check documentation}
#| eval: false
?carData::OBrienKaiser
```

A multi-variate linear model is fit to the data.

```{r fit multivairate model}
mod_OK <- lm(cbind(pre.1, pre.2, pre.3, pre.4, pre.5,
                   post.1, post.2, post.3, post.4, post.5,
                   fup.1, fup.2, fup.3, fup.4, fup.5) ~  treatment * gender,
             data = OBrienKaiser)
```


To test a specific hypothesis, an L- and M-Matrix can be specified. Here the null hypothesis is: "*on average, there is no difference between the pre and the follow-up measures*". Considering the contrasts for `treatment` and ` gender`, the corresponding L-matrix is:

```{r L-matrix}
#| tbl-cap: "L-matrix to test a within effect on average participants"
L <- rbind(c(1, 0, 0, 0, 0, 0))
dimnames(L) <- list("average", rownames(coef(mod_OK)))
knitr::kable(L)
```

The M-matrix can be defined as follows:

```{r M-matrix}
#| tbl-cap: "M-matrix to test a difference between pre and follow-up measures"
M <- cbind(rep(c(-1/5, 0, 1/5), each = 5))
dimnames(M) <- list(colnames(coef(mod_OK)), "fu-pre")
knitr::kable(M)
```

The `linearHypothesis()` function from the `car` package prints the test results, but the returned object does not contain the printed values. Yet, invisible a `linearHypothesis.mlm` object is returned.


```{r linear hypothesis test}
test <- car::linearHypothesis(mod_OK, L, P = M, test = "Wilks")
test
```


In order to extract the test results, the `extract_results()` function from the `aRm` package can be used:

```{r extract results}
results <- extract_results(test)
knitr::kable(results, digits = c(1, 3, 1, 0, 0, 5))
```

As the returned object is a `data.frame`, extracting results is straightforward.

```{r}
p <- results$'p-value'
```


## Example - Anova

It is also possible to specify the design of the repeated measures:

```{r}
idata <- expand.grid(measure = factor(1:5), 
                     time = factor(c("pre", "post", "fup")))
contrasts(idata$measure) <- contr.sum(5)
contrasts(idata$time) <- contr.sum(3)

anova <- car::Anova(mod_OK, 
                    idata = idata, 
                    idesign = ~ measure * time, 
                    type = "III")


```

In order to extract the test results, the `extract_results()` function from the `aRm` package can be used:

```{r extract results}
results <- extract_results(anova)
knitr::kable(results, digits = c(1, 3, 1, 0, 0, 5))
```



